Icicle: a single-pass query language

### 2015-12-05 2016icicle/section/02-Source.tex
changes

We formally define the grammar of the language in figure~\ref{fig:source:grammar}.
Note that general application is not allowed: arguments can only be applied to primitives and named functions.
Similarly, functions and primitives must be fully applied.
This, combined with the lack of lambda construction, means that higher order functions are outlawed.
Outlawing higher order functions is somewhat unsatisfying, but simplifies the typing rules and makes it easier to support our performance guarantees.

>>>>

We formally define the grammar of the language in figure~\ref{fig:source:grammar}.
Note that general application is not allowed: arguments can only be applied to primitives and named functions.
Similarly, functions and primitives must be fully applied.
This, combined with the lack of lambda construction, means that higher order functions are outlawed.
While a first-order language (one lacking higher order functions) offers somewhat less abstraction, it simplifies the typing rules and makes it easier to support our performance guarantees.

### 2015-12-05 2016icicle/section/03-Types.tex
changes

%%% Expressions
With function application already defined, typechecking expressions in figure~\ref{fig:source:type:exp} is quite simple.
The judgment $\Typecheck{\Gamma}{x}{\tau}$ means that under a context $\Gamma$, the expression $x$ has type $\tau$.
Rules (TcVar) and (TcPrim) simply look up the variable or primitive in the environment.
For (TcApp) we simply use the boxed function application judgment from figure~\ref{fig:source:type:wrap}.

\TODO{I think we need a box/cast rule for going from @Pure@ to @Element@ etc}

%%% Contexts
Figure~\ref{fig:source:type:ctx} shows the typing rules for contexts.
The judgment $\Typecheck{\Gamma}{c}{\tau}$ means that under a context $\Gamma$, the context $c$ has type $\tau$.
For simplicity, we reuse the same judgment for expressions and contexts.

>>>>

%%% Expressions
With function application already defined, typechecking expressions in figure~\ref{fig:source:type:ctx} becomes quite standard.
The judgment $\Typecheck{\Gamma}{x}{\tau}$ means that under a context $\Gamma$, the expression $x$ has type and modality $\tau$.
Being a first-order language, the typing rule returns a first-order type rather than a function type.

Rules (TcVar) and (TcPrim) simply look up the variable or primitive in the environment.

For (TcApp) we simply use the boxed function application judgment from figure~\ref{fig:source:type:wrap}.

\TODO{I think we need a box/cast rule for going from @Pure@ to @Element@ etc}
For simplicity, we reuse the same judgment for expressions and contexts.
